### 2009-04-07 talk/icooolps2009-dotnet/benchmarks.tex
Some statments rephrased, minor corrections

The design goal of the language is to be very simple (the interpreter of the
full language consists of about 600 lines of RPython code) but to still have
the typical properties of dynamic languages that make them hard to
compile. TLC is implemented with a small interpreter that interprets a custom
bytecode instruction set. Since our main interest is in the runtime
performance of the interpreter, we did not implement the parser nor the
bytecode compiler, but only the interpreter itself.

@@

Moreover, for each benchmark we also show the time taken by running the
equivalent program written in C\#.\footnote{The sources for both TLC and C\#
  programs are available at:

  \texttt{http://codespeak.net/~antocuni/tlc-benchmarks/}}

@@

Table \ref{tab:factorial-fibo} shows the seconds spent to calculate
the factorial and Fibonacci for various $n$.  As we can see, for small values
of $n$ the time spent running the JIT compiler is much higher than the time
spent to simply interpret the program.  This is an expected result
which, however, can be improved once we will have time
to optimize compilation and not only the generated code.

@@

As already discussed, our generated JIT does not compile the whole function at
once. Instead, it compiles and executes code chunk by chunk, waiting until it
knows enough informations to generate highly efficient code.  In particular,
at the time it emits the code for the inner loop it exactly knows the
type of \lstinline{obj}, thus it can remove the overhead of dynamic dispatch
and inline the method call.  Moreover, since \lstinline{obj} never escapes the
function, the allocation is avoided and its field \lstinline{value} is stored
as a local variable.  As a result, the generated code turns out to be a simple loop
doing additions in-place.

>>>>

The design goal of the language is to be very simple (the interpreter of the
full language consists of about 600 lines of RPython code) but to still have
the typical properties of dynamic languages that make them hard to
compile. TLC is implemented with a small interpreter that interprets a custom
bytecode instruction set. Since our main interest is in the runtime
performance of the interpreter, we did not implement the whole language, but just its virtual machine.

@@

Moreover, for each benchmark we also show the time taken by running the
equivalent program written in C\#.\footnote{The sources for both TLC and C\#
  programs are available at:
  \texttt{http://codespeak.net/~antocuni/tlc-benchmarks/}}

@@

Table \ref{tab:factorial-fibo} shows the seconds spent to calculate
the factorial and Fibonacci for various $n$.  As we can see, for small values
of $n$ the time spent running the JIT compiler is much higher than the time
spent to simply interpret the program.  This is an expected result
which, however, can be improved, since so far no effort has been direct to
enhance the performance of the compiler itself.

@@

As already discussed, our generated JIT does not compile the whole function at
once. Instead, it compiles and executes code chunk by chunk, waiting until it
knows enough information to generate highly efficient code.  In particular,
at the time it emits the code for the inner loop it exactly knows the
type of \lstinline{obj}, thus it can remove the overhead of dynamic dispatch
and inline the method call.  Moreover, since \lstinline{obj} never escapes the
function, the allocation is avoided and its field \lstinline{value} is stored
as a local variable.  As a result, the generated code turns out to be a simple loop
doing additions in-place.
